---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>",
  fig.path = file.path("man", "figures", "README-"),
  out.width = "100%"
)
```

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License](https://img.shields.io/badge/license-GPL--3-blue.svg?style=flat)](https://www.gnu.org/licenses/gpl-3.0.html)
[![Release](https://img.shields.io/github/release/inbo/inbovegtypering.svg)](https://github.com/inbo/inbovegtypering/releases)
![GitHub](https://img.shields.io/github/license/inbo/inbovegtypering)
[![R build status](https://github.com/inbo/inbovegtypering/workflows/check%20package%20on%20main/badge.svg)](https://github.com/inbo/inbovegtypering/actions)
![r-universe name](https://inbo.r-universe.dev/badges/:name?color=c04384)
![r-universe package](https://inbo.r-universe.dev/badges/inbovegtypering)
[![Codecov test coverage](https://codecov.io/gh/inbo/inbovegtypering/branch/main/graph/badge.svg)](https://app.codecov.io/gh/inbo/inbovegtypering?branch=main)
![GitHub code size in bytes](https://img.shields.io/github/languages/code-size/inbo/inbovegtypering.svg)
![GitHub repo size](https://img.shields.io/github/repo-size/inbo/inbovegtypering.svg)
<!-- badges: end -->

# inbovegtypering

The goal of inbovegtypering is to classify vegetation relevés (plots) into predefined vegetation types based on the methods described by van Tongeren et al. (2008).

It provides a single function, classify_vegetation(), to calculate various indices:

- Composite Distance (CoD) (Eq. 9)
- 2ln(Likelihood) (Eq. 4)
- Weirdness (Eq. 5)
- Incompleteness (Eq. 6)
- Modified Euclidean Distance (MED) (Eq. 8)

## Installation

You can install the development version from GitHub with:

```{r}
# install.packages("remotes")
remotes::install_github("inbo/inbovegtypering")
```

## Workflow

The classification process relies on three key pieces of data:

Relevé Data: Your own plot data. This must contain:

- RecordingGivid: A unique ID for each plot.
- LayerCode: The vegetation layer (e.g., "K" for herb layer).
- PctValue: The percentage cover (0-100) for each species.
- species_number: A list-column containing the internal species ID(s) that match the synoptic table. The read_inboveg_recording() function is designed to generate this.

Synoptic Table:

- The reference data. This table defines the vegetation types. It must contain:
- syntaxonCode: The unique ID for the vegetation type.
- speciesNumber: The internal species ID.
- frequency: The frequency of the species in the type (0-100).
- mean_if_present: The average cover of the species when it is present (0-100).

Species List: 

A helper table (used internally by read_inboveg_recording) to link species names and GBIF keys to the speciesNumber in the synoptic table.

## Example 1: Basic Classification (with Dummy Data)

This example shows the core classification workflow using self-contained dummy data.

```{r}
library(inbovegtypering)
library(tibble)
library(dplyr)

# --- 1. Prepare Data ---

# A) Create a dummy Synoptic Table
# We have two types: "Forest_A" (rich in species 101, 102)
# and "Grassland_B" (rich in species 103, 104)
dummy_synoptic_table <- tibble::tribble(
  ~syntaxonCode, ~speciesNumber, ~frequency, ~mean_if_present,
  "Forest_A",     101,            80,         15,
  "Forest_A",     102,            70,         10,
  "Forest_A",     103,            10,         5,
  "Forest_A",     104,            5,          1,
  "Grassland_B",  101,            5,          2,
  "Grassland_B",  102,            10,         2,
  "Grassland_B",  103,            90,         25,
  "Grassland_B",  104,            85,         20
)

# B) Create a dummy Relevé (Plot)
# This plot has high cover of species 103 & 104, so it should match "Grassland_B".
# The `species_number` column MUST be a list.
dummy_releve <- tibble::tribble(
  ~RecordingGivid, ~LayerCode, ~PctValue, ~species_number,
  "Plot_1",        "K",        20,        list(103L),
  "Plot_1",        "K",        15,        list(104L),
  "Plot_1",        "K",        2,         list(101L)
)

# --- 2. Classify the Relevé ---

# Use the main function 'classify_vegetation()'
# We will test two different methods

# Method: Composite Distance (CoD)
cls_cod <- classify_vegetation(
  data = dummy_releve,
  synoptics = dummy_synoptic_table,
  method = "cod"
)

# Method: -2ln(Likelihood)
cls_llk <- classify_vegetation(
  data = dummy_releve,
  synoptics = dummy_synoptic_table,
  method = "likelihood"
)

# --- 3. Inspect the Results ---

# A) Print the classification object
# It shows the top matches (lower values are better)
print(cls_cod)

# B) Get a statistical summary
summary(cls_cod)

# C) Plot the classification rose
# The longest spoke is the best match
plot(cls_cod, types = c("Forest_A", "Grassland_B"))

# Plot the 'likelihood' results
plot(cls_llk, types = c("Forest_A", "Grassland_B"))
```

## Example 2: Real Workflow (with Database)

This is the intended workflow for connecting to INBO databases. This example will not run without a database connection and credentials.

```
library(inbovegtypering)
library(dplyr)

# --- 1. Connect to Databases ---
# This requires credentials and connection to the INBO network
con_inboveg <- connect_db_inboveg()
con_taxa <- connect_db_taxonomy()

# --- 2. Read Relevé Data ---
# This function gets the plot data from Cydonia AND
# links it to the taxonomy database to create the 'species_number' column.
my_releve <- read_inboveg_recording(
  con_inboveg = con_inboveg,
  con_taxa = con_taxa,
  survey = "MILKLIM_Heischraal2012",
  code = "IV2012081611384756"
)

# --- 3. Classify Relevé ---
# We can use synoptics = "default" to load the synoptic table
# included with the inbovegtypering package.
classification <- classify_vegetation(
  data = my_releve,
  synoptics = "default",
  method = "cod"
)

# --- 4. Inspect Results ---
print(classification)
summary(classification)
plot(classification, types = 12) # Plot the top 12 matches
```
