---
output: github_document
title: "Inbovegtypering package"
---

# Inbovegtypering package

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License](https://img.shields.io/badge/license-GPL--3-blue.svg?style=flat)](https://www.gnu.org/licenses/gpl-3.0.html)
[![Release](https://img.shields.io/github/release/inbo/inbovegtypering.svg)](https://github.com/inbo/inbovegtypering/releases)
![GitHub](https://img.shields.io/github/license/inbo/inbovegtypering)
[![R build status](https://github.com/inbo/inbovegtypering/workflows/check%20package%20on%20main/badge.svg)](https://github.com/inbo/inbovegtypering/actions)
![r-universe name](https://inbo.r-universe.dev/badges/:name?color=c04384)
![r-universe package](https://inbo.r-universe.dev/badges/inbovegtypering)
[![Codecov test coverage](https://codecov.io/gh/inbo/inbovegtypering/branch/main/graph/badge.svg)](https://app.codecov.io/gh/inbo/inbovegtypering?branch=main)
![GitHub code size in bytes](https://img.shields.io/github/languages/code-size/inbo/inbovegtypering.svg)


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>",
  fig.path = file.path("man", "figures", "README-"),
  out.width = "100%"
)
```


The goal of inbovegtypering is to classify vegetation relevés (plots) into predefined vegetation types (syntaxa) based on the methods described by van Tongeren et al. (2008). It implements the ASSOCIA algorithm for supervised clustering.

## Installation

You can install the development version of inbovegtypering from GitHub with:

```{r install}
# install.packages("devtools")
devtools::install_github("inbo/inbovegtypering")

```


## Methodology

The core function classify_vegetation() calculates similarity indices between your vegetation recordings and a reference synoptic table.

The indices include:

- **Composite Distance (CoD):** The primary classification metric combining qualitative and quantitative data.

- **(un)Likelihood (-2ln):** A qualitative distance measure based on species presence/absence.

- **Weirdness:** Measures the presence of unexpected species.

- **Incompleteness:** Measures the absence of expected species.

- **Modified Euclidean Distance (MED):** A quantitative distance measure based on species abundance (cover).

**Note:** For all these indices, a **lower value** indicates a **better** match.

## Workflow

The classification process requires two main inputs:

1. **Recordings Data:** Your plot data containing species cover values.

1. **Synoptic Table:** A reference table defining the vegetation types (syntaxa) with species frequencies and characteristic cover.


### Example: Basic Classification

This example uses the dummy datasets included in the package (example_recordings and example_synoptics).


```{r examplesnippet}
library(inbovegtypering)
library(dplyr)

# 1. Load Example Data
# --------------------
data("dummy_recordings")
data("dummy_synoptics")

# Inspect the recordings (5 plots)
head(dummy_recordings)

# 2. Run Classification
# ---------------------
# We classify the recordings against the provided synoptic table.
# By default, this calculates the Composite Distance (CoD) and all sub-indices.
classification_results <- classify_vegetation(
  data = dummy_recordings,
  synoptics = dummy_synoptics,
  layer = "K" # Analyze the herb layer
)

# 3. Inspect Results
# ------------------

# Print the result object (shows top matches for each recording)
print(classification_results)

# Get a detailed summary for the first recording
# Shows top 5 matches sorted by CoD
summary(classification_results, top_n = 5)

# 4. Visualization
# ----------------

# A) Classification Rose Plot
# Visualizes the "distance" to the best matching syntaxa.
# Longer spokes = Better Match.
# We plot the results for REC-01 and REC-02 (which are very similar)
plot(classification_results, 
     indextype = "cod", 
     types = 10, 
     ncol = 2, nrow = 1)

# B) Classification Bar Chart
# Shows the distance scores directly (Shorter bars = Better Match)
# Useful for comparing the exact metric values.
barplot(classification_results, 
        indextype = "cod", 
        types = 10,
        ncol = 2, nrow = 1)
```


## Using Default Synoptics

If the package includes a default synoptic table in inst/extdata/synoptic_table.csv, you can use the shorthand "default" argument:


```{r defaultsynoptics}
results <- classify_vegetation(
  data = my_recordings,
  synoptics = "default"
)
```


## References

van Tongeren, O., Gremmen, N., & Hennekens, S. (2008). Assignment of relevés to pre-defined classes by supervised clustering of plant communities using a new composite index. Journal of Vegetation Science, 19(4), 525-536.


